---
title: "World Debt"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    theme: cosmo
---

```{r, error = TRUE, message=FALSE, warning=FALSE, include=FALSE}

rm(list=ls())
library(flexdashboard)
library(dplyr)
library(ggplot2)
library(magrittr)
library(rvest)
library(tidyverse)
library(leaflet)
library(RColorBrewer)
library(plotly)
library(readr)
library(tidyquant)
library(urbnmapr)
library(tmap)
library(sf)
library(stringr)
library(tigris)
library(rgdal)
library(WDI)
library(highcharter)
library(rworldmap)
library(tidyr)
library(igraph)
library(networkD3)
```

Overview
===================================== 
### Project Topic: Debt Levels 

This is our final project for QMSS 5063: Data Visualization (Columbia University, Spring 2021).   
Members: Luke Artola, Samantha Li, Yun Yee Tan (Group K)

With the rise of COVID-19, governments around the world have drastically increased their amount of debt. Within the developed countries, the US' debt-to-GDP ratio has now become 128% and Japan has skyrocketed to 234%, which are staggering numbers. The developing countries have been hit with an even harder blow. Given this renewed interest and concern over debt levels, we decided to gear our project towards providing information that might be of interest to those who work on fiscal and economic development policy.

We will be examining government debt from a few different levels:  

1. First, we will look at national debt within the US and how debt levels vary across states.  

2. Next, we will zoom out on a global level to explore the geographical distribution of debt as well as its relationship with the Human Development Index.  

3. Finally, we will explore international debt flows and the network between debtor and creditor nations.

Our findings demonstrate that due to the continual increase in national debt in developing and developed countries, policy analysts should keep an eye on the ongoing situation and mitigate potential negative implications sovereign debt can have through investing in programs contributing to economic growth and understanding their nations debt flows and network.

Global Debt
===================================== 

### The Financial Effects of Covid
The surprise COVID-19 pandemic has left a lasting impression on the world, one in which will not be forgotten anytime soon. One of the effects of the pandemic has been a unprecedented surge in sovereign debt. As of October 2020 the Debt to GDP ratio of advanced economies was a staggering 123.9% with emerging markets resulting in a 62.5%. Developed economies have not seen this level of debt since the second great war. The major difference between then and now is governments feel much more comfortable holding larger balance sheets and do not seem to have the political will power to make the necessary fiscal adjustments needed to reduce it. Within the past year governments increased their borrowing by 19.5 trillion dollars alone  Source(https://www.bloomberg.com/graphics/2021-coronavirus-global-debt/#:~:text=The%20Covid%2D19%20Pandemic%20Has%20Added%20%2419.5%20Trillion%20to%20Global%20Debt)


In this section we will explore the surge in sovereign debt due to the COVID-19 pandemic and in particular focus on the fiscal health of the United States.


### Covid Debt Surge 2020
Within the past year alone countries have seen double digit increases in sovereign debt with countries such as Canada, US, and Japan seeing at least a 25% increase in their debt commitments. This is rather worrying for some countries in particular as highlighted previously with the United States having over 100% debt to GDP and japan being well over 200% debt to gdp. 

### Covid Debt Surge 2020

```{r}
data <- read.csv("covid.csv")
data <- data %>%
  rename(Country = "X")


p <- ggplot(data = data, aes( x = Country, y = Debt_Increase)) +
  geom_bar(stat = "identity", fill = "dodgerblue1") +
  geom_text(aes(label=Debt_Increase), vjust = .5, hjust = 2, color="black", size=5) +
  scale_x_discrete(limits=c("China","France", "Spain", "United Kingdom", "Japan", "United States", "Canada")) +
  ggtitle("Debt to GDP Change in 2020") +
  xlab("Country") +
  labs(caption = "Bloomberg, IFF") +
  ylab("Percent Increase") +
  coord_flip() +
  theme_classic()

p +  theme(axis.text.x = element_text(face = "bold", colour = "black")) +
  theme(axis.text.y = element_text(face = "bold", colour = "black")) 

```

### US National Debt
Historically prior to 2000 every major increase in the United States debt to GDP was a result of war or economic downturn which subsequently reduced after a short period of time. This is not the case anymore as we can see from the graph the debt has been steadily increasing over the past 20 years with two major spikes due to the financial crisis of 08 and now the COVID-19 pandemic. The current debt to GDP ratio of the United States is 128% which is the largest the United States has ever had. The Congressional budget office using the assumption of ceteris paribus regarding new legislation and recessions projects that within the next 30 the US debt will reach over 200% with a little over 10% annual deficit. Source(CBO)

### US National Debt to GDP Ratio

```{r}
d = read.csv("debt.csv")
us_debt = read.csv("federal_debt_trends.csv") 

usa = us_debt
usa = select(us_debt, year, federal_debt, gdp,debt_gdp)

# dot-and-line chart


usa_chart <- ggplot(usa, aes(x = year, y = debt_gdp,
                                             text = paste0("<b>Year: </b>", year,"<br>",
                                                           "<b>Debt: </b>", federal_debt," dollars<br>",
                                                           "<b>GDP: </b> $", gdp, " dollars<br>",
                                                          "<b>Debt GDP ratio: </b> ", debt_gdp, "(ratio)"),
                                             group = 1)) +
  xlab("Year") +
  ylab("Dollars ($)") +
  theme_minimal(base_size = 10, base_family = "Georgia") +
  geom_point() +
  geom_line()

usa_chart_interactive <- ggplotly(usa_chart, tooltip = "text") %>% 
  config(displayModeBar = FALSE) %>%
  layout(hoverlabel = list(bgcolor = "white",
                           font = list(family = "Georgia")))

usa_chart_interactive

```

### US State Debt
Many of the services provided to people on a daily basis is a result of state and local programs that are not funded by the federal government. Given that it is also important to keep an eye on the fiscal health of state governments. As it is rare for a state to default on their debt it is not unheard of as Arkansas in the 40's ultimately defaulted on their loan obligations. Based on the map it seems that many states are doing relatively well managing their balance sheets but there are a few states that should continued to be looked at closely. Illinois and New Jersey in particular currently is at 468% and 440% debt to GDP. As it is not an immediate concern given another economic downturn these states and others with elevated levels of debt will find it difficult to fiance future budgets without receiving government assistance which is not always guaranteed.

### US State Debt

```{r}
data <- read.csv("state_2020.csv") 
data_1 <- data %>%
  select(State, Debt.Ratio..Total.Debts.Total.Assets) %>%
  rename("Debt_Level" = "Debt.Ratio..Total.Debts.Total.Assets")

#Debt was recorded as a character, converted to numeric
data_1$Debt_Level <- str_remove_all(string = data_1$Debt_Level, pattern = "%")
data_1$Debt_Level <- as.numeric(data_1$Debt_Level)



# removing DC as it is a district
states_sf <- get_urbn_map("states", sf = TRUE)
states_sf <- states_sf %>%
  filter(state_abbv != "DC")

states_sf <- states_sf %>%
  rename(State = "state_name")

state_debt <- left_join(data_1, states_sf, by = "State")
#head(state_debt)

state_debt <- st_as_sf(state_debt)


position <- c("center", "top")
tm_shape(state_debt) + 
  tm_polygons(col = "Debt_Level", title = "Debt Level", palette = "Purples", breaks = c(0,25,50,75, 100, 150,200,300,400, 500)) +
  tm_text(text = "state_abbv", size = .75, remove.overlap = TRUE) +
  tm_legend(scale = .80) + 
  tm_layout(title = "State Debt Levels 2020 (%)", title.position = position) 


```

HDI 
===================================== 

### Human Development Index's (HDI) Relationship to Debt

According to the UN, http://hdr.undp.org/en/hdp-covid  the Human Development Index (HDI) examines three important criteria of economic development (life expectancy, education and income levels) and uses this to create an overall score between 0 and 1.

A 2013 study by the World Bank found that if the debt-to-GDP ratio exceeds 77% for an extended period, it slows economic growth (HDI). Every percentage point of debt above this level costs the country 0.017 percentage points in economic growth. Emerging markets are even more sensitive to debt-to-GDP ratios. In such markets, each additional percentage point of debt above 64% will slow growth by 0.02 percentage points each year. (Source: https://elibrary.worldbank.org/doi/abs/10.1596/1813-9450-5391)

According to another literature by Zaghdoudi 2018, the optimal debt level is 41.7775%. Below this debt threshold, external debt has a positive effect on human development. Any 1% increase in the external debt ratio induces an increase in the HDI of 0.02%. However, above the debt threshold, external debt becomes detrimental to human development since HDI decreases by 0.01% when external debt ratio increases by 1%. (Source: https://ideas.repec.org/a/ebl/ecbull/eb-18-00571.html)

This idea is further bolstered by a study Amani Alzahrani conducted in 2018, proving that Government debt affects HDI and FDI (Foreign Direct Investment) because it influences all aspects of investment. In particular, it reduces public investment. "As government debt increases, the government spends more of its budgets on interest costs, and such a move crowds out public investment by affecting local and foreign investors. In the US, for example, the CBO estimated in 2017 that the interest costs of government debt is likely to reach $5.2 trillion. In other words, the interest charges will triple the current program used by the government to run the national economic needs." This to a potential hypothesis that to reverse this debt increase, a method of doing so would be to invest in programs that increase public investment (contributing to higher HDI such as education, etc) which will help reduce debt levels in the long run. 

In our research, we have chosen a variety of countries based on a mix of developed and developing countries due to their interesting backgrounds to examine both their debt:gdp ratio and the hdi levels for a holistic comparison and review (to see if we can gain interesting insights). Especially given COVID-19, where the crisis is hitting hard on all of human development’s constitutive elements, we wanted to see if we can advise current policies based on historical data. Comprehensive debt relief and restructuring is essential to ensure a recovery from COVID that is both inclusive and sustainable. This is because investment in these programs will ultimately decrease debt, and therefore why we should be looking at HDI as a metric to evaluate how we can reduce our debt by investing into programs contributing to economic growth. 

### Understanding HDI around the world

The UN Human Development Report is released every few year with updated figures. Given that the most complete dataset was in 2015, we have chosen to examine the HDI of the world to gain a better understanding about the status of a country than the Per Capita Income method by World Bank. The map demonstrates that North America and European countries generally have a higher HDI than the rest of the world. 

### HDI 2015 Chloropleth World Map

```{r}
hdi = read.csv("human-development-index-escosura.csv")
hdiworld = hdi
#hdi

hdiworld <- hdiworld %>%
  filter(Year == 2015) %>%
  select(Entity, Historical.Index.of.Human.Development..Prados.de.la.Escosura.)  
#hdiworld
colnames(hdiworld)[colnames(hdiworld) == "Historical.Index.of.Human.Development..Prados.de.la.Escosura."] <- "HDI"
# Load required R packages


# Set highcharter options
options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))

# Load the world Map data
data(worldgeojson, package = "highcharter")

#worldgeojson
hdiworld2 <-hdi
hdiworld2 <- hdiworld2 %>%
  filter(Year == 2015) %>%
  select(Code, Historical.Index.of.Human.Development..Prados.de.la.Escosura.)  

#hdiworld2
colnames(hdiworld2)[colnames(hdiworld2) == "Historical.Index.of.Human.Development..Prados.de.la.Escosura."] <- "HDI"

hc2 <- highchart() %>%
  hc_add_series_map(
    worldgeojson, hdiworld2, value = "HDI", joinBy = c('iso3','Code'),
    name = "HumanDevelopmentIndex"
    )  %>% 
  hc_colorAxis(stops = color_stops()) %>% 
  hc_title(text = "World Map") %>% 
  hc_subtitle(text = "HDI in 2015")

hc2
```


### Understanding HDI around the world
We've chosen 5 different countries with varying HDI levels to examine the trend between HDI and debt:GDP ratio. This research examines Argentine, China, India, Peru, and the United States because of their interesting economic statuses. We see that the US takes the lead in HDI, followed by the rest. 

### HDI Over Time by Country

```{r}
# create interactive line chart
graph3 = hdi

graph3 <- graph3 %>%
  select(Year, Entity, Historical.Index.of.Human.Development..Prados.de.la.Escosura.)  

colnames(graph3)[colnames(graph3) == "Historical.Index.of.Human.Development..Prados.de.la.Escosura."] <- "HDI"
#graph3

plotdata <- spread(graph3, Entity, HDI)

h <- highchart() %>% 
  hc_xAxis(categories = plotdata$Year) %>% 
  hc_add_series(name = "Argentina", 
                data = plotdata$Argentina) %>% 
  hc_add_series(name = "China", 
                data = plotdata$China) %>%
  hc_add_series(name = "India", 
                data = plotdata$India) %>%
  hc_add_series(name = "Peru", 
                data = plotdata$Peru) %>%
  hc_add_series(name = "United States of America", 
                data = plotdata$"United States") 

h <- h %>%
  hc_title(text = "HDI by Country",
           margin = 20, 
           align = "left",
           style = list(color = "steelblue")) %>% 
  hc_subtitle(text = "1870 to 2015",
              align = "left",
              style = list(color = "#2b908f", 
                           fontWeight = "bold")) %>% 
  hc_credits(enabled = TRUE, # add credits
             text = "United Nations HDI") %>% 
  hc_legend(align = "left", 
            verticalAlign = "top",
            layout = "vertical", 
            x = 0, 
            y = 100) %>%
  hc_tooltip(crosshairs = TRUE, 
             backgroundColor = "#FCFFC5",
             shared = TRUE, 
             borderWidth = 4) %>% 
  hc_exporting(enabled = TRUE)

h


```


```{r, warning=FALSE, message=FALSE}
world_debt = read_csv("world_debt - Sheet1.csv")
world_debt[world_debt == "no data"] <- NA
df_transpose = as.data.frame(t(world_debt))
names(df_transpose) <- lapply(df_transpose[1, ], as.character)
df_transpose <- df_transpose[-1,] 
```

### Understanding the Debt:GDP ratio by country
"The debt-to-GDP ratio is the metric comparing a country's public debt to its gross domestic product (GDP). By comparing what a country owes with what it produces, the debt-to-GDP ratio reliably indicates that particular country's ability to pay back its debts." (https://www.investopedia.com/terms/d/debtgdpratio.asp)

By knowing the debt:gdp ratio with the HDI, we can understand how each country is managing their debt and if it extended beyond the 77% for developed countries and the 64% for developing countries as indicated by the World Bank to verify if it slows economic growth. 

### Debt:GDP Ratio by Country 

```{r}
world_debt_gdp_ratio = df_transpose
# generate graph
#Finding debt over time for:
#Argentina
#China
#India
#Peru
#US
world_debt_gdp_ratio = world_debt_gdp_ratio[2:48,] %>% mutate_if(is.character,as.double)

world_debt_gdp_ratio$Year = row.names(world_debt_gdp_ratio)
world_debt_gdp_ratio <- world_debt_gdp_ratio[-(1:10),] 

w <- highchart() %>% 
  hc_xAxis(categories = world_debt_gdp_ratio$Year, na.RM=TRUE)  %>% 
  hc_add_series(name = "Argentina", 
                data = world_debt_gdp_ratio$Argentina, na.RM=TRUE,
                color = "#D55400") %>% 
  hc_add_series(name = "China", 
                data = world_debt_gdp_ratio$"China, People's Republic of",na.RM=TRUE,
                color = "#227FBB") %>% 
  hc_add_series(name = "India", 
                data = world_debt_gdp_ratio$India,na.RM=TRUE,
                color = "#1ECE6D") %>% 
  hc_add_series(name = "Peru", 
                data = world_debt_gdp_ratio$Peru,na.RM=TRUE,
                color = "#F2C500") %>% 
  hc_add_series(name = "United States of America", 
                data = world_debt_gdp_ratio$"United States",na.RM=TRUE,
                color= '#2B3E51') 

# customize interactive line chart
w <- w %>%
  hc_title(text = "Debt:GDP Ratio by Country",
           margin = 20, 
           align = "left",
           style = list(color = "steelblue")) %>% 
  hc_subtitle(text = "1990 to 2025",
              align = "left",
              style = list(color = "#2b908f", 
                           fontWeight = "bold")) %>% 
  hc_credits(enabled = TRUE, # add credits
             text = "International Monetary Fund") %>% 
  hc_legend(align = "left", 
            verticalAlign = "top",
            layout = "vertical", 
            x = 0, 
            y = 100) %>%
  hc_tooltip(crosshairs = TRUE, 
             backgroundColor = "#FCFFC5",
             shared = TRUE, 
             borderWidth = 4) %>% 
  hc_exporting(enabled = TRUE)
w



```

### A Holistic View on Advanced Economies versus Developing Countries Debt:GDP Ratio Over Time
The graph below demonstrates that even over the past few years, debt:gdp ratio has gradually increased over time for both advanced and developing countries. This verifies the fact that advanced economies have a higher debt:gdp ratio and developing countries have lower debt:gdp ratio, and therefore, developing countries are more sensitive to the impacts of incurring debt (thus lower threshold is needed to pass to slow economic growth, which is consistent with the HDI trends above.)

### Advanced Economies versus Developing Countries Debt:GDP Ratio Over Time

```{r}

world_debt_gdp_ratio <- world_debt_gdp_ratio[-(1:10),] 

w2 <- highchart() %>% 
  hc_xAxis(categories = world_debt_gdp_ratio$Year) %>% 
  hc_add_series(name = "Advanced Economies", 
                data = world_debt_gdp_ratio$`Advanced economies`) %>%
  hc_add_series(name = "Developing Countries", 
                data = world_debt_gdp_ratio$`Emerging market and developing economies`) 

# customize interactive line chart
w2 <- w2 %>%
  hc_title(text = "Debt:GDP Ratio by Type",
           margin = 20, 
           align = "left",
           style = list(color = "steelblue")) %>% 
  hc_subtitle(text = "2000 to 2025",
              align = "left",
              style = list(color = "#2b908f", 
                           fontWeight = "bold")) %>% 
  hc_credits(enabled = TRUE, # add credits
             text = "world_debt_gdp_ratio") %>% 
  hc_legend(align = "left", 
            verticalAlign = "top",
            layout = "vertical", 
            x = 0, 
            y = 100) %>%
  hc_tooltip(crosshairs = TRUE, 
             backgroundColor = "#FCFFC5",
             shared = TRUE, 
             borderWidth = 4) %>% 
  hc_exporting(enabled = TRUE)

w2




```

International Debt Flows
===================================== 

```{R, warning=FALSE, message=FALSE}
dssi <- read_csv("dssi_2020.csv")

dssi <- dssi %>% select(-`Series Code`) %>% gather("Year", "Value", 6:9)

dssi <- dssi %>% anti_join(dssi[c(29785, 29786, 29787, 29788, 29789, 59574, 59575, 59576, 59577, 59578, 89363, 89364, 89365, 89366, 89367, 119152, 119153, 119154, 119155, 119156),])

dssi <- dssi %>% spread(`Series Name`, `Value`)

colnames(dssi) <- c("Debtor", "Debtor_Code", "Creditor", "Creditor_Code", "Year", "Debt_Forgiveness", "Debt_Stocks")

dssi$Debt_Forgiveness <- as.numeric(dssi$Debt_Forgiveness)
dssi$Debt_Stocks <- as.numeric(dssi$Debt_Stocks)
dssi$Year <- as.factor(dssi$Year)

dssi[is.na(dssi)] <- 0

dssi$Debtor <- stringr::str_trim(dssi$Debtor, side = "both")
dssi$Creditor <- stringr::str_trim(dssi$Creditor, side = "both")

dssi <- dssi %>% mutate(Debtor_Region = case_when(
  Debtor %in% c("Angola", "Benin", "Burkina Faso", "Burundi", "Cabo Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Congo, Dem. Rep.", "Congo, Rep.", "Cote d'Ivoire", "Djibouti", "Ethiopia", "Gambia, The", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mozambique", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Sierra Leone", "Somalia", "Tanzania", "Togo", "Uganda", "Zambia") ~ "Africa",
  Debtor %in% c("Bangladesh", "Bhutan", "Cambodia", "Kyrgyz Republic", "Lao PDR", "Maldives", "Mongolia", "Myanmar", "Nepal", "Pakistan", "Tajikistan", "Timor-Leste", "Uzbekistan") ~ "Asia",
  Debtor %in% c("Afghanistan", "Yemen, Rep.") ~ "Middle East",
  Debtor %in% c("Guyana", "Haiti", "Honduras", "Nicaragua") ~ "South & Central America",
  Debtor %in% c("Dominica", "Grenada", "St. Lucia", "St. Vincent and the Grenadines") ~ "The Caribbeans",
  Debtor %in% c("Fiji", "Papua New Guinea", "Samoa", "Solomon Islands", "Tonga", "Vanuatu") ~ "Pacific Islands",
  Debtor %in% c("Kosovo", "Moldova") ~ "Europe"
))


debt_series <- read_csv("debt_series.csv")

debt_series <- debt_series[1:68, c(1, 7:26)]
colnames(debt_series) <- c("Debtor", 2000:2019)
debt_series <- debt_series %>% gather(key = "Year", value = "Debt_Stocks", 2:21)

debt_series$Debt_Stocks <- as.numeric(debt_series$Debt_Stocks)
debt_series$Year <- as.numeric(debt_series$Year)

debt_series[which(is.na(debt_series$Debt_Stocks)), "Debt_Stocks"] <- 0

debt_series <- debt_series %>% mutate(Debtor_Region = case_when(
  Debtor %in% c("Angola", "Benin", "Burkina Faso", "Burundi", "Cabo Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Congo, Dem. Rep.", "Congo, Rep.", "Cote d'Ivoire", "Djibouti", "Ethiopia", "Gambia, The", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mozambique", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Sierra Leone", "Somalia", "Tanzania", "Togo", "Uganda", "Zambia") ~ "Africa",
  Debtor %in% c("Bangladesh", "Bhutan", "Cambodia", "Kyrgyz Republic", "Lao PDR", "Maldives", "Mongolia", "Myanmar", "Nepal", "Pakistan", "Tajikistan", "Timor-Leste", "Uzbekistan") ~ "Asia",
  Debtor %in% c("Afghanistan", "Yemen, Rep.") ~ "Middle East",
  Debtor %in% c("Guyana", "Haiti", "Honduras", "Nicaragua") ~ "South & Central America",
  Debtor %in% c("Dominica", "Grenada", "St. Lucia", "St. Vincent and the Grenadines") ~ "The Caribbeans",
  Debtor %in% c("Fiji", "Papua New Guinea", "Samoa", "Solomon Islands", "Tonga", "Vanuatu") ~ "Pacific Islands",
  Debtor %in% c("Kosovo", "Moldova") ~ "Europe"
))

selected_creditors <- read_csv("selected_creditors.csv")

selected_creditors <- selected_creditors %>% select(-`Series Code`)
colnames(selected_creditors) <- c("Debtor", "Debtor_Code", "Creditor", "Creditor_Code", "Series", 2000:2019)

selected_creditors <- selected_creditors %>% gather("year", "value", 6:25)

selected_creditors[which(selected_creditors$Series == "External debt stocks, total (DOD, current US$)"), "Series"] <- "Debt_Stocks"
selected_creditors[which(selected_creditors$Series == "Debt forgiveness or reduction (current US$)"), "Series"] <- "Debt_Forgiveness"

selected_creditors <- selected_creditors %>% filter(!is.na(Series))

selected_creditors <- spread(selected_creditors, Series, value)


selected_creditors$Debt_Forgiveness <- as.numeric(selected_creditors$Debt_Forgiveness)
selected_creditors$Debt_Stocks <- as.numeric(selected_creditors$Debt_Stocks)
selected_creditors$year <- as.numeric(selected_creditors$year)

selected_creditors[is.na(selected_creditors)] <- 0

selected_creditors$Debtor <- stringr::str_trim(selected_creditors$Debtor, side = "both")
selected_creditors$Creditor <- stringr::str_trim(selected_creditors$Creditor, side = "both")

selected_creditors <- selected_creditors %>% mutate(Debtor_Region = case_when(
  Debtor %in% c("Angola", "Benin", "Burkina Faso", "Burundi", "Cabo Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Congo, Dem. Rep.", "Congo, Rep.", "Cote d'Ivoire", "Djibouti", "Ethiopia", "Gambia, The", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mozambique", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Sierra Leone", "Somalia", "Tanzania", "Togo", "Uganda", "Zambia") ~ "Africa",
  Debtor %in% c("Bangladesh", "Bhutan", "Cambodia", "Kyrgyz Republic", "Lao PDR", "Maldives", "Mongolia", "Myanmar", "Nepal", "Pakistan", "Tajikistan", "Timor-Leste", "Uzbekistan") ~ "Asia",
  Debtor %in% c("Afghanistan", "Yemen, Rep.") ~ "Middle East",
  Debtor %in% c("Guyana", "Haiti", "Honduras", "Nicaragua") ~ "South & Central America",
  Debtor %in% c("Dominica", "Grenada", "St. Lucia", "St. Vincent and the Grenadines") ~ "The Caribbeans",
  Debtor %in% c("Fiji", "Papua New Guinea", "Samoa", "Solomon Islands", "Tonga", "Vanuatu") ~ "Pacific Islands",
  Debtor %in% c("Kosovo", "Moldova") ~ "Europe"
))

```


### International Debt Flows

Given the major blow that COVID-19 has dealt to the world’s poorest countries, in 2020, the World Bank and the International Monetary Fund established the Debt Service Suspension Initiative with the G-20. Under this initiative, which has been extended through December 2021, 73 countries are eligible for a temporary suspension of debt-service payments owed to their official bilateral creditors. 

Part of this initiative includes greater debt transparency. For the first time last year, the World Bank released the breakdown of individual creditor countries. Previously, such data was only available at a generalized “creditor region” level. This newly-released database provides us with a lot of insight on the vast network of debt flows between debtor and creditor nations. 

This section will explore international debt flows and the relationships between debtors and creditors. We will first look at the overall patterns of debt levels over time. Then, using the newly released data from the World Bank, we will examine the breakdown of creditors, including countries and multilateral organizations. Finally, we will explore the entire network of debtor nations and their creditors. The data shows us that global debt has increased drastically in the past two decades, and this increase is most significant for Asian countries. On the creditor side, the most striking trend comes from the dramatic increase in lending from China. The international debt landscape is expected to continue evolving over time and governments and policy analysts should continue to keep a close eye on these dynamics.

Data source: International Debt Statistics (The World Bank, 2021)
https://databank.worldbank.org/source/international-debt-statistics:-dssi

Note: All debt levels in the following visualizations are denominated in current US$


### Total debt per region (2000-2019)

Firstly, I wanted to get an overview of how debt levels have changed over time. Here we can see that debtor countries from Asia and Africa have experienced a significant rise in total debt levels over the past two decades, compared to other regions. 

### Total debt per region (2000-2019)

```{r, warning=FALSE, message=FALSE}
ggplotly(debt_series %>% group_by(Debtor_Region, Year) %>% summarize(total_debt_bil = sum(Debt_Stocks)/10^9 %>% round(digits = 2), mean_debt_bil = mean(Debt_Stocks)/10^9 %>% round(digits = 2)) %>% ggplot() + geom_line(aes(x = Year, y = total_debt_bil, color = Debtor_Region)) + labs(y = "Total debt (billions of US$)", title = "Total debt per region over time") + theme_minimal())
```


### Mean debt per region (2000-2019)

To account for the different group sizes (e.g. half of the debtor nations are from Africa, which would disproportionately inflate the total amount of debt in Africa), it might make more sense to compare average debt levels across regions instead. Here, the picture has changed - mean debt levels in Asia have risen at a much greater rate than the other regions. All regions have experienced an increase in mean debt level, with the exception of the Caribbean nations whose mean debt levels have remained relatively constant. 

### Mean debt per region (2000-2019)

```{r, warning=FALSE, message=FALSE}
ggplotly(debt_series %>% group_by(Debtor_Region, Year) %>% summarize(total_debt_bil = sum(Debt_Stocks)/10^9 %>% round(digits = 2), mean_debt_bil = mean(Debt_Stocks)/10^9 %>% round(digits = 2)) %>% ggplot() + geom_line(aes(x = Year, y = mean_debt_bil, color = Debtor_Region)) + labs(y = "Mean debt (billions of US$)", title = "Mean debt per region over time") + theme_minimal())

```

### Total lending among top creditors (2000-2019)

Next, I wanted to visualize how the breakdown of individual creditors has changed over time. The line graph here shows that total lending has increased significantly for organizations like the World Bank and Asian Development Bank. In terms of creditor countries, we see a drastic increase in lending from China, while the US and France have decreased their total lending. 

### Total lending among top creditors (2000-2019)

```{r, warning=FALSE, message=FALSE}
creditors_years <- selected_creditors %>% group_by(Creditor, year) %>% summarize(total_lending_bil = sum(Debt_Stocks)/10^9)


ggplotly(ggplot(creditors_years) + geom_line(aes(x = year, y = total_lending_bil, color = Creditor)) + labs(x = "Year", y = "Total lending (billions of US$)", title = "Total lending over time (top creditors)") + theme_minimal()) 

```

### Breakdown of top creditors (2000 and 2019)
This pattern can also be visualized through the next two pie charts, which compares the breakdown of top 20 creditors for 2000 vs 2019. The most significant change comes from China, which was represented by a small slice of the pie in 2000 and has jumped to be the top creditor nation in 2019. This is no surprise given China’s increasing investments in developing countries such as through the Belt and Road Initiative and other debt diplomacy arrangements. In contrast, countries which took up a larger portion of the pie in 2000 such as France. Russia and the US are now represented by a much smaller slice. 

### Breakdown of top creditors (2000)

```{r, warning=FALSE, message=FALSE}
cols <- RColorBrewer::brewer.pal(8, "Dark2")
debt_2000 <- dssi %>% filter(Year == 2000)
debt_2019 <- dssi %>% filter(Year == 2019)

top_creditors_2000 <- debt_2000 %>% filter(!Creditor %in% c("World", "Multiple Lenders", "Other Multiple Lenders", "Other Multilaterals", "Bondholders", "Other Bilateral")) %>% group_by(Creditor) %>% summarize(total_debt_bil = sum(Debt_Stocks)/10^9 %>% round(digits = 3)) %>% arrange(desc(total_debt_bil)) %>% head(n = 20)

top_creditors_2019 <- debt_2019 %>% filter(!Creditor %in% c("World", "Multiple Lenders", "Other Multiple Lenders", "Other Multilaterals", "Bondholders", "Other Bilateral")) %>% group_by(Creditor) %>% summarize(total_debt_bil = sum(Debt_Stocks)/10^9 %>% round(digits = 3)) %>% arrange(desc(total_debt_bil)) %>% head(n = 20) 

top_creditors_2000 %>% hchart("pie", hcaes(x = Creditor, y = round(total_debt_bil, digits = 1)), name = "Total Lending (billions of US$)") %>% hc_title(text = "Top creditors in 2000")  %>% hc_caption(text = "Data: The World Bank (2021)") %>% hc_colors(cols)

```

### Breakdown of top creditors (2019)

```{r, warning=FALSE, message=FALSE}
top_creditors_2019 %>% hchart("pie", hcaes(x = Creditor, y = round(total_debt_bil, digits = 1)), name = "Total Lending (billions of US$)") %>% hc_title(text = "Top creditors in 2019")  %>% hc_caption(text = "Data: The World Bank (2021)") %>% hc_colors(cols)

```

### Debtor-Creditor Network (2019)

Lastly, this visualization shows the network between all debtors and creditors. The size of the nodes is proportional to the total debt levels (in 2019) for that particular country or organization, and the width of the arrows is proportional to the amount of debt flows between two particular nodes. I have also tried to color the nodes in accordance to their geographical region (e.g. countries in Asia/Oceania/Middle East are shades of red and orange, while countries in the Americas/Caribbeans are shades of blue). 

China, for instance, lends to countries in Africa, Asia and Oceania, whereas regional banks are more targeted, such as the Inter-American Development Bank which only lends to nations in South and Central America. Interestingly, Pakistan borrows mostly from European countries. Click on the individual nodes for more information on total debt/lending levels. 

### Debtor-Creditor Network (2019)

```{r, warning=FALSE, message=FALSE}
links_2019 <- debt_2019 %>% select(Debtor, Creditor, Debt_Stocks) %>% filter(Debt_Stocks != 0) %>% filter(!Creditor %in% c("Other Bilateral", "Other Multiple Lenders", "Other Multilaterals", "World", "Multiple Lenders", "Bondholders"))

links_2019 <- links_2019[, c(2, 1, 3)]

links_2019$Debt_Stocks <- links_2019$Debt_Stocks/10^9 

debtors_2019 <- links_2019 %>% group_by(Debtor) %>% summarize(size = sum(Debt_Stocks))
debtors_2019$group <- "Debtor"
debtors_2019 <- debtors_2019 %>% rename(name = Debtor)

creditors_2019 <- links_2019 %>% group_by(Creditor) %>% summarize(size = sum(Debt_Stocks))
creditors_2019$group <- "Creditor"
creditors_2019 <- creditors_2019 %>% rename(name = Creditor)

nodes_2019 <- rbind(debtors_2019, creditors_2019)

nodes_2019 <- as.data.frame(nodes_2019)
links_2019 <- as.data.frame(links_2019)

nodes_2019$id <- 1:nrow(nodes_2019)

nodes_2019 <- nodes_2019 %>% mutate(region = case_when(
  name %in% c("Angola", "Benin", "Burkina Faso", "Burundi", "Cabo Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Congo, Dem. Rep.", "Congo, Rep.", "Cote d'Ivoire", "Djibouti", "Ethiopia", "Gambia, The", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mozambique", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Sierra Leone", "Somalia", "Tanzania", "Togo", "Uganda", "Zambia", "Cote D`Ivoire, Republic Of", "Libya", "Mauritius", "South Africa") ~ "Africa",
  name %in% c("Bangladesh", "Bhutan", "Cambodia", "Kyrgyz Republic", "Lao PDR", "Maldives", "Mongolia", "Myanmar", "Nepal", "Pakistan", "Tajikistan", "Timor-Leste", "Uzbekistan", "China", "Hong Kong", "India", "Japan", "Korea, Republic Of", "Kuwait", "Malaysia", "Singapore", "Sri Lanka", "Thailand") ~ "Asia",
  name %in% c("Afghanistan", "Yemen, Rep.", "Bahrain", "Egypt", "Iran, Islamic Republic Of", "Israel", "Saudi Arabia", "United Arab Emirates") ~ "Middle East",
  name %in% c("Guyana", "Haiti", "Honduras", "Nicaragua", "Brazil", "Venezuela, Republic Bolivarian") ~ "South & Central America",
  name %in% c("Dominica", "Grenada", "St. Lucia", "St. Vincent and the Grenadines", "Bahamas", "Barbados", "St. Kitts And Nevis", "Trinidad & Tobago", "St. Vincent & The Grenadines") ~ "The Caribbeans",
  name %in% c("Canada", "United States") ~ "North America", 
  name %in% c("Fiji", "Papua New Guinea", "Samoa", "Solomon Islands", "Tonga", "Vanuatu", "Australia") ~ "Oceania",
  name %in% c("Kosovo", "Moldova", "Argentina", "Austria", "Belarus", "Belgium", "Czech Republic", "Denmark", "Finland", "France", "Germany, Fed.Rep. Of", "Greece", "Hungary", "Ireland", "Italy", "Netherlands", "Norway", "Poland", "Portugal", "Russian Federation", "Serbia", "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey", "United Kingdom") ~ "Europe",
  name %in% c("African Dev. Bank", "Asian Dev. Bank", "Inter-American Dev. Bank", "International Monetary Fund", "World Bank-IBRD", "World Bank-IDA") ~ "Organizations"
))


links_2019 <- left_join(links_2019, nodes_2019 %>% select(name, id), by = c("Creditor" = "name"))
links_2019 <- left_join(links_2019, nodes_2019 %>% select(name, id), by = c("Debtor" = "name"))

links_2019 <- links_2019 %>% rename(source = id.x, target = id.y)

links_2019$source <- links_2019$source - 1
links_2019$target <- links_2019$target - 1


```

```{r, warning=FALSE, message=FALSE}
script <- 'alert("Name: " + d.name + "\\n" + "Status: " + d.status + "\\n" + "Region: " + d.group + "\\n" + "Total debt/lending (billions of US$): " + d.size);'

network_2019 <- forceNetwork(Links = links_2019, 
                   Nodes = nodes_2019, 
                   Source = "source",
                   Target = "target", 
                   Value = "Debt_Stocks", 
                   NodeID = "name",
                   Nodesize = "size",
                   Group = "region", 
             fontSize = 20,
             linkDistance = 80,
                   opacity = 0.9, zoom = FALSE, legend = TRUE, arrows = TRUE, height = 500, width = 1000, bounded = TRUE, clickAction = script,
             colourScale = JS('d3.scaleOrdinal()
           .range(["#ce8000", "#619e8d", "#fa8072", "#584a91", "#ff748c", "6897bb", "a75572", "c7c9c7", "0a75ad"]);'))



network_2019$x$nodes$status <- nodes_2019$group
network_2019$x$nodes$size <- nodes_2019$size %>% round(digits = 3)


## Click on the node to see info popup
network_2019
```


